---
title: "PSY4210 Lab Report - 2023 - Part 2"
author: "33777594 -- Farhanul Majumder"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: 
    toc: true
---

# Set up

```{r packages, message=FALSE}

library(haven)
library(data.table)
library(JWileymisc)
library(ggplot2)
library(ggpubr)
library(visreg)

# Add other packages you need here
library(tufte)
library(mice)
library(VIM)
library(lme4)
library(lmerTest)
library(multilevelTools)
library(emmeans)

```

# Loading Data

```{r data setup}
## read in the dataset
data(aces_daily)
d <- as.data.table(aces_daily)

## between person data, no missing
davg <- na.omit(d[, .(
 Female = factor(na.omit(Female)[1], levels = 0:1),
 Age = na.omit(Age)[1],
 STRESS = mean(STRESS, na.rm = TRUE),
 PosAff = mean(PosAff, na.rm = TRUE),
 NegAff = mean(NegAff, na.rm = TRUE)),
 by = UserID])

## create missing data
davgmiss <- copy(davg)
davgmiss[STRESS < 1, NegAff := NA]
davgmiss[STRESS > 4, PosAff := NA]

## random missingness on age
set.seed(1234)
davgmiss[, Age := ifelse(rbinom(
 .N, size = 1, prob = .1) == 1,
 NA, Age)]

## drop unneeded variables to make analysis easier
davgmiss[, UserID := NULL]
```

# 3. Missing Data

```{r}

## multiply impute data
mi.1 <- mice(
  davgmiss,
  m = 5,   maxit = 30,
  defaultMethod = c("norm", "logreg", "polyreg", "polr"),
  seed = 1234, printFlag = FALSE)

## convergence plot
plot(mi.1, PosAff + NegAff + Age ~ .it | .ms)

## linear regression with imputed dataset
mi.reg <- with(mi.1, lm(STRESS ~ NegAff + Age))

mi.reg #print

pool.r.squared(mi.reg) #pooled results with R^2 values

```

## Missing Data Write up:

The dataset "davgmiss" contained missing data. Multiple imputations were performed to recover the data. 5 imputed datasets were generated, each with 30 iterations. Visually analysing the convergance plot generated revealed that the data was converging. A linear regression with the imputed dataset was performed, with stress as the outcome and negative affect and age as predictors. Results were pooled with an R-Squared value of 0.608, 95% CI = [0.49, 0.70].

# 4. Linear Mixed Models 1

```{r fig.width = 7, fig.height = 10}

ri.m <- lmer(PosAff ~ 1 + (1 | UserID),
            data = d,
            REML = TRUE)

summary(ri.m)


## calculate intraclass correlation coefficient
iccMixed("PosAff", id = "UserID", data = d)

## run model diagnostic
plot(modelDiagnostics(ri.m), ncol = 2, nrow = 2, ask = FALSE)

## Confidence interval of the model
confint(ri.m, method = "profile", oldNames = FALSE)


```

## Linear Mixed Models 1 Write up:

An intercept only linear mixed model was fit to 6399 positive affect scores from 191 people. The intraclass correlation coefficient of positive affect, was 0.5431, indicating that about 54% of the total variance in positive affect was between people and the other 46% was within person due to daily fluctuations. Visual diagnostics showed that positive affect was normally distributed, hence normality assumption appear to be met. Assumptions of homogeneity of variance were met however some extreme values were found (refer to QQ deviates plot). The fixed effect intercept showed that average positive affect was 2.679, 95% CI =[2.56, 2.79]. There were however individual differences in positive affect with standard deviation for random intercept being 0.7930.

# 5. Linear Mixed Models 2

```{r fig.width = 7, fig.height = 10}
## create a between and within STRESS variable
d[, c("bSTRESS", "wSTRESS") := meanDeviations(STRESS), by = UserID]

## linear mixed model with bSTRESS and wSTRESS predicting NegAff
m1 <- lmer(PosAff ~ bSTRESS + wSTRESS + (1 + wSTRESS | UserID), data = d)

## run model diagnostics
m1.diag <- modelDiagnostics(m1, ev.perc = .005)
plot(m1.diag, ncol = 2, nrow = 3, ask = FALSE)

## identify extreme values
m1.diag$extremeValues

## creating new dataset with extremevalues removed and rerun model
dnoev <- d[-unique(m1.diag$extremeValues$Index)]
m1noev <- lmer(PosAff ~ bSTRESS + wSTRESS + (1 + wSTRESS | UserID), data = dnoev)
summary(m1noev) #final model summary

## confidence interval for new model
m1noev.ci <- confint(m1noev, oldNames = FALSE)
m1noev.ci

##rerun diagnostics
m1noev.diag <- modelDiagnostics(m1noev, ev.perc = .005)
plot(m1noev.diag, ncol = 2, nrow = 3, ask=FALSE)

```

## Linear Mixed Models 2 Write up:

A linear mixed model was fit to 6249 positive affect scores from 189 people. The fixed effect intercept revealed that the estimated average stress was 3.1824 [95% CI = 2.99, 3.37], when between-persons and within-person stress was zero. There were individual differences, with standard deviation of 0.7362. There was a significant fixed effect of average stress on positive affect, such that one unit higher average stress was associated with -0.2241 [95% CI = -0.29, -0.16] lower positive affect (p \<0.001). On the days where individuals were high on stress, people were expected to have -0.1595 lower positive affect, 95% CI = [-0.17, -0.15] (p \<0.001).

# 6. Interactions and Moderation for LMMs

```{r}
## Generate LMM
m2 <- lmer(STRESS ~ SUPPORT + COPEPrb + SUPPORT:COPEPrb + (1 | UserID), data = d)
APAStyler(modelTest(m2),
          pcontrol = list(digits = 3, stars = FALSE, includeP = TRUE,
                          includeSign = TRUE, dropLeadingZero = TRUE))

## finding M and SD of COPEPrb to generate simple slope for interaction
egltable(c("COPEPrb"), data = d[!duplicated(UserID)])

## test simple slope
mem <- emtrends(m2, var = "SUPPORT",
                at = list(COPEPrb = c(2.37-0.66, 2.37+0.66)),
                lmer.df = "satterthwaite")
summary(mem, infer = TRUE)

## Create interaction graph
visreg(m2, xvar = "SUPPORT", 
       by = "COPEPrb", overlay=TRUE, 
       breaks = c(2.37-0.66, 2.37+0.66), 
       partial = FALSE, rug = FALSE, gg=TRUE, 
       xlab = "Social Support", 
       ylab = "Predicted Stress Rating") + 
  scale_color_manual(values = c("1.71" = "black", "3.03" = "grey70")) +
  theme_pubr() +
  guides(colour = FALSE, fill = FALSE) +
  annotate(geom = "text", x = 6, y = 2.9, 
           label = "High Problem-Focused Coping: b = -0.24 [ -0.30, -0.18], p <.0001",
           angle = -29) + 
  annotate(geom = "text", x = 5.5, y = 2.3, 
           label = "Low Problem-Focused Coping: b = -0.2 [-0.24, -0.15], p <.0001",
           angle = -24)
  




```

## Interaction and Moderation for LMMs Write up:

A linear mixed model using restricted maximum likelihood was used to test whether the association of social support on daily stress rating is moderated by problem-focused coping scores. All predictors were included as fixed effects and a random intercept by participant was included.

The social support x problem-focused coping interaction was not statistically significant (-0.04,[95% CI = -0.08, 0.01], p = .122) , which indicated that the relationship between social support and stress did not vary by problem-focused coping. If the interaction was significant, a simple slope analysis could reveal that social support would lead to a decrease in stress at both high and low problem-focused coping.
