---
title: "PSY4210 Lab Report - 2023 - Part 1" 
author: "33777594 – Farhanul Bari Majumder" 
date: "`r Sys.Date()`"
output: 
  tufte::tufte_html: 
    toc: true
---

# 1. Generalized Linear Models 1

```{r}

library(haven)
library(data.table)
library(JWileymisc)
library(ggplot2)
library(ggpubr)
library(visreg)


## read in data
db <- as.data.table(read_sav("[2021] PSY4210 BL.sav")) # baseline data

```

## Run a Linear Regression

Fit **one** moderated, multiple linear regression. Pick from:

-   Outcome: Anxiety at baseline
-   Predictors: `sex` and any other variable of your choosing
-   Make sure that you include the interaction between your predictor and `sex`.

Note: you will need to score Anxiety from the STAI. For a refresher, see the content for Data Visualization 1.

```{r glm1reg}

#creating variable "anxiety"
db[, anxiety := rowMeans(.SD, na.rm = TRUE) * 4,
   .SDcols = c("STAI1", "STAI2", "STAI3", "STAI4")]

# turning sex into a factor
db[, sex := factor( 
  sex,
  levels = c(1,2),
  labels = c("male", "female"))]

db [, BFI_C1 := 6- BFI_C1r] #reverse scoring
db [, conscientiousness := rowMeans(.SD, na.rm = TRUE),
    .SDcols = c("BFI_C1", "BFI_C2")] # creating variable "conscientiousness"


## put your regression model code using lm() here

m <- lm(anxiety ~ sex * conscientiousness, data = db)


```

## Linear Regression Diagnostics

Conduct model diagnostics on your regression model. Where applicable, apply appropriate transformations and/or exclude outliers.

```{r glm1diagnostics}

## put your regression diagnostics code including plots using modelDiagnostics() here
md <- modelDiagnostics(m, ev.perc = 0.005)

plot(md, ncol = 2, ask = FALSE)

```

## Linear Regression Summary

Create a summary of your **final** model, after any relevant transformations or extreme values have been addressed, where applicable.

-   use `modelTest()` and `APAStyler()` to get a nice result.

Create a graph to help visualize your results using `visreg()`. Note, you may want to adjust the `fig.width` and `fig.height` options to this `R` chunk to make the graph have the ratios you like.

```{r glm1plot, fig.width = 8, fig.height = 6}

## put your code to summarize the final model using modelTest() here
summary(m)
confint(m) # calculating confidence interval
mtest <- modelTest(m)
knitr::kable(APAStyler(mtest)) # clean presentation

## put your code to visualize your regression model using visreg() here
visreg(m, xvar = "conscientiousness", by = "sex", 
       partial = FALSE, rug = FALSE,
       overlay = TRUE, gg = TRUE) +
  theme_pubr() +
  ggtitle("Multiple regression for anxiety - separate lines for each sex group")

```

In the last subheading (Linear Regression Write Up) for GLM1, briefly summarize your analysis steps, assumption checks, any changes performed, and interpret the interaction and regression coefficients, referencing your figure visualizing the result where appropriate. The total length should be about one paragraph and about 150 words.

## Linear Regression Write Up

A multiple linear regression model (with interaction) was created with outcome variable anxiety and predictors conscientiousness and sex group. Total anxiety score was generated from summing the STAI questions. The first conscientiousness question (labelled "BFI_C1r") was reverse scored. Assumptions of normality were met with normally distributed errors (as seen in the residual distribution curve) and homoscedasticity. There was a statistically significant association between anxiety and conscientiousness, controlling for sex group. As visualised in the negative slope of regression figure, a one unit higher anxiety score was associated with a -1.87 [95% CI -3.63 - -0.10] lower conscientiousness score, p \< 0.05. Anxiety and sex group association was non-significant (-3.37, p = .366). The interaction between predictors sex and conscientiousness was non-significant (1.45, p = .127). Overall, conscientiousness and sex group explained 8.72% of the variance in self-esteem.

# 2. Generalized Linear Models 2

1.  Conduct one multiple logistic regression predicting high selfesteem "SelfesteemHigh" as created in the GLM2 lecture content **from two predictors of your choosing.**

2.  Create a graph of the predicted probabilities of being in the SelfesteemHigh group by each of your two predictors. It should parallel the predicted probability plot in section 4.4 of the GLM2.rmd 

3.  Report and interpret the odds ratios, 95% confidence intervals for each predictor and present the average marginal effect for each predictor to interpret the results on the probability scale.

## Question 1: GLM High Self Esteem

```{r}

#create self-esteem from LSE questions
db[, selfesteem:= rowMeans(.SD, na.rm = TRUE),
.SDcols = c("LSE1", "LSE2", "LSE3", "LSE4")]
db[, SelfesteemHigh := as.integer(selfesteem >= 3.75)] #create "SelfesteemHigh" variable

#create extraversion variable (predictor) from BFI_E questions
db [, BFI_E1 := 6- BFI_E1r] # reverse scoring
db[, extraversion := rowMeans(.SD, na.rm = TRUE), 
     .SDcols = c("BFI_E1", "BFI_E2")] # creating variable "extraversion"

#create agreeableness variable (predictor) from BFI_A questions
db [, BFI_A2 := 6- BFI_A2r] # reverse scoring
db [, agreeableness := rowMeans(.SD, na.rm = TRUE),
    .SDcols = c("BFI_A1", "BFI_A2")] #creating variable "agreeableness"

mlog <- glm(SelfesteemHigh ~ extraversion + agreeableness,
             data = db, family = binomial())
summary(mlog)
```

## Question 2: High self esteem probability graph

```{r}

#predicted probability graph for SelfesteemHigh
visreg(mlog, xvar = "extraversion",
       by = "agreeableness", breaks = c(3, 4.5),
       scale = "response", overlay = TRUE,
       partial = FALSE, rug = FALSE, gg = TRUE) +
  ylab("predicted probability of being high self esteem") +
  theme_pubr()
```

## Question 3: Odds Ratio and Marginal Effects

```{r}

# Calculate Odds Ratio
exp(coef(mlog))

# Calculate Confidence intervals
exp(confint(mlog))

#calculate marginal effect, using h = .01
h <- .01

# making original dataset to interpret with predictors, excluding missing values
originaldata <- db[!is.na(extraversion),
                   .(extraversion, agreeableness)]

# dataset with increased extraversion only
IncreasedExtraversion <- db[!is.na(extraversion), 
                   .(extraversion = extraversion + h, agreeableness)]

# dataset with increased agreeableableness only
IncreasedAgreeableness <- db[!is.na(extraversion), 
                   .(extraversion, agreeableness = agreeableness + h)]

# calculating predicted probabilities for original, extraversion and agreeableness
originaldata$Prob <- predict(mlog, newdata = originaldata,
                             type = "response")

IncreasedExtraversion$Prob <- predict(mlog, newdata = IncreasedExtraversion,
                             type = "response")

IncreasedAgreeableness$Prob <- predict(mlog, newdata = IncreasedAgreeableness,
                             type = "response")

# calculating the difference in probabilities per unit
ex.diffprob <- (IncreasedExtraversion$Prob - originaldata$Prob) / h
ag.diffprob <- (IncreasedAgreeableness$Prob - originaldata$Prob) / h

# calculating the average marginal effects
mean(ex.diffprob) # extraversion
mean(ag.diffprob) # agreeableness

```

## Odds Ratio and Marginal Effect Write up

A multiple logistic regression predicting high self esteem from extraversion and agreeableness showed that for each additional unit of extraversion, participants have 1.86 times the odds of having high self-esteem [95% CI 1.33 - 2.66], p \< .001. For each additional unit of agreeableness, participants have 1.76 times the odds of having high self-esteem [95% CI 1.17 - 2.73], p \< .01. When holding agreeableness constant, increasing extraversion by one unit on average is predicted to result in 0.13 change (increase) in the probability of having high self-esteem. Alternatively, when holding extraversion constant, increasing agreeableness by one unit on average is predicted to results in 0.12 change (increase) in the probability of having high self-esteem.
